---
title: "ISQA 521 Final Project"
author: "Skye Gilbreth, Jordan Hilton"
date: "May 2, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lessR)
library(ggplot2)
library(tidyverse)
library(reshape2)
```

# Introduction

We're beginner chess players, and we'd like to improve our game a little. We have a data set of 20,000 chess games played on the website lichess, including a computer-identified opening used in each game. The goal of this project is to analyze this set of games to find out more about what openings Iwe should study, by looking at the performance of each opening using this data.

The set is from this kaggle:

https://www.kaggle.com/datasnaek/chess

# Data Load

Let's load the data and take a look at it. We've dropped out some unnecessary columns, including the largest one which is a list of moves for each game.

```{r loaddata}
chessdata <- read.csv("data/chessdata.csv")
chessdata <- chessdata[,-c(1,3,4,13,14,16)]
head(chessdata)
```

Let's go over our available fields quickly. We have:

* rated, whether or not the game was rated
* turns, how many turns were played 
* victory_status, how the game ended (by timeout, resignation, draw, or mate)
* winner, the outcome of the game (white wins, black wins, or a draw)
* increment_code, a description of the time control for the game (15+3 means that each player started with 15 minutes on the clock and had a 3 second increment for each move)
* white_id and black_id, the ids of the players
* white_rating and black_rating, the ELO ratings of the players for the category of game they're playing
* opening_name, the computer's description of what opening was played

We're now going to create two additional fields by mutation; one, using regular expressions, consolidates categories of openings (so "Queen's Pawn Game: Han Solo Variant | The Coolest Line" will be grouped in with all the other Queen's Pawn games), and one calculates an average of the two players' ratings.

```{r fixnames}
chessdata<-mutate(chessdata, shortname=gsub(":.*","",chessdata$opening_name))
chessdata<-transform(chessdata, shortname=gsub("\\|.*","",chessdata$shortname))
chessdata<-mutate(chessdata, avgrating=(white_rating+black_rating)/2)
head(chessdata)
```

Who's the highest rated player we have?

```{r highestratedplayer}
head(sort(chessdata$white_rating, decreasing=TRUE), 1)
filter(chessdata, white_rating==2700)
```

The answer is "justicebot", long may he reign, playing the Caro-Kann.

# Exploratory charts

Let's take a look at what we have in our new "shortnames" field in terms of number of games for each opening:

```{r listofnames}
length(table(chessdata$shortname))
```

180 is too many openings to neatly chart! Let's filter our table down to the top 15 openings and then take a look at some bar charts.


```{r filter}
top15openings<-head(names(sort(table(chessdata$shortname), decreasing=TRUE)),15)
filtereddata<-filter(chessdata, shortname %in% top15openings)
```

```{r}
#Export altered data to a CSV file to be used by the Shiny app.
write.csv(filtereddata, "data/filtereddata.csv")
```

Now let's start with a basic bar chart just to see the relative frequency of these top 15 openings.

```{r basicbarchart}
BarChart(shortname, data=filtereddata, horiz=TRUE, quiet=TRUE, xlab="Opening", sort="-", ylab="Count of Games")
```

The Sicilian Defense (1. e4 c6) is by far the most popular, with an even distribution after that. We're keeping an eye on the Italian Game (1. e4 e5 2. Nf3 Nc6 3. Bc4) since that's Jordan's favorite opening, and we're pleased to see it in 4th place.

As an excuse to switch to histograms, how about the ratings of the players in this data set?

```{r ratinghisto}
Histogram(white_rating, data=filtereddata, fill="lightgray", quiet=TRUE, xlab="White Player's Rating", ylab="Count of Games")
```

This looks like a normal distribution with a slight skew to the left. The default rating is 1500, so there are probably a lot of games by new players at or immediately under 1500. Note the slight uptick at the very top of the range - a buildup of high-rated players.

Are the ratings of the black players any different?
```{r ratinghisto2}
Histogram(black_rating, data=filtereddata, fill="black", quiet=TRUE, xlab="Black Player's Rating", ylab="Count of Games")
```



\pagebreak

No, the distribution of ratings for the black players are identical; that's as expected. How long is a normal game?

```{r turnshisto}
Histogram(turns, data=filtereddata, bin.width=15, quiet=TRUE, xlab="Number of Turns", ylab="Count of Games")
```

It looks like a normal game takes around 50 turns, but there's a long tail of games that go all the way out to 350 moves- those are probably complicated endgames with little advantage where somebody is refusing to resign or accept a draw.


Let's take a look at the 4 most frequent openings (couldn't leave out the Italian!). Do any of them get noticeably more frequent in games by higher-rated players, which might indicate that they're just a better opening?

```{r top4openingratingline}
top4<-head(names(sort(table(chessdata$shortname), decreasing=TRUE)),4)
filtereddata2<-filter(chessdata, shortname %in% top4)
ggplot(filtereddata2, aes(x=white_rating, color=shortname))+geom_line(stat="bin", binwidth=25)+ylim(0,100)
```

How about the relative win rates of each opening?

```{r winratebarchart}
BarChart(shortname, by=winner, data=filtereddata, horiz=TRUE, ylab="Count of Games", stat.x="proportion", xlab="Opening Name",quiet=TRUE)
```



# Best Opening for Me

---highest winrate by rating

```{r winratebyrating}
filtereddata$ratingbin<-cut(filtereddata$white_rating, seq(0,3500,500), labels=c("0","1","2","3","4","5","6"))
##BarChart(ratingbin, by=winner, by1=shortname, data=filtereddata, horiz=TRUE, ylab="Count of Games", stat.x="proportion", xlab="Opening Name",quiet=TRUE)
ggplot(filtereddata, aes(x = ratingbin, fill=winner)) +  
        geom_bar(aes(y = (..count..)/sum(..count..)),position="fill") + 
        scale_y_continuous()+facet_wrap(vars(shortname)) +labs(x="Binned White Rating", y="Proportion of Games")
```

Note that each opening seems to improve with rating- this is a methodological error because we've used the white player's rating while ignoring the black player's rating, so with a higher rating just comes a higher win rate! Let's switch to the average rating instead:

```{r winratebyrating2}
filtereddata$ratingbin<-cut(filtereddata$avgrating, seq(0,3500,500), labels=c("0","1","2","3","4","5","6"))
##BarChart(ratingbin, by=winner, by1=shortname, data=filtereddata, horiz=TRUE, ylab="Count of Games", stat.x="proportion", xlab="Opening Name",quiet=TRUE)
ggplot(filtereddata, aes(x = ratingbin, fill=winner)) +  
        geom_bar(aes(y = (..count..)/sum(..count..)),position="fill") + 
        scale_y_continuous()+facet_wrap(vars(shortname)) +labs(x="Binned White Rating", y="Proportion of Games")
```

Here we get to some interesting results. Picking some in particular: the Scotch game and the King's Pawn games are famously sharp, so it doesn't surprise me that they perform well at high and low average rating but poorly in middling ratings. The Four Knights is supposed to be bad, so it makes sense that its performance gets worse at higher ratings.


# Shiny Interactive

# Conclusion

We should definitely study the Sicilian, since it is by far the most common opening and it performs well at high ratings. In the meantime, though, sticking with the Italian looks like it should work out just fine; it performs decently at all rating levels and doesn't get too drawish at the high rating levels.

### TODO:

1. Shiny interactive - a barchart of frequency and winrate for top 15 openings, with a slider for your rating? Needs both a barchart and histogram. Might just include the URL for final document for proff. to access.

2. Fix winratebyrating chunk - Jordan

3. Add more explanatory text and analysis

4. Add conclusion -(would like it to be based off of winrate by rating)

5. Add following graphics: 

--- draw rate by opening and rating
  maybe an overlapping hs? barplot? Should we include something with factors or bubble plots?

6. Look over homeworks to see if there are any chart types we think are missing