---
title: "ISQA 521 Final Project"
author: "Jordan Hilton"
date: "May 2, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lessR)
library(ggplot2)
library(tidyverse)
library(reshape2)
```

# Introduction

I'm a beginner chess player, and I'd like to improve my game a little. I have a data set of 20,000 chess games played on the website lichess, including a computer-identified opening used in each game. The goal of this project is to analyze this set of games to find out more about what openings I should study, by looking at the performance of each opening using this data.

The set is from this kaggle:

https://www.kaggle.com/datasnaek/chess

# Data Load

Let's load the data and take a look at it. I've dropped out some unnecessary columns, including the largest one which is a list of moves for each game.

```{r loaddata}
chessdata <- read.csv("chessdata.csv")
chessdata <- chessdata[,-c(1,3,4,13,14,16)]
head(chessdata)
```

Let's go over our available fields quickly. We have:

* rated, whether or not the game was rated
* turns, how many turns were played 
* victory_status, how the game ended (by timeout, resignation, draw, or mate)
* winner, the outcome of the game (white wins, black wins, or a draw)
* increment_code, a description of the time control for the game (15+3 means that each player started with 15 minutes on the clock and had a 3 second increment for each move)
* white_id and black_id, the ids of the players
* white_rating and black_rating, the ELO ratings of the players for the category of game they're playing
* opening_name, the computer's description of what opening was played

```{r fixnames}
chessdata<-mutate(chessdata, shortname=gsub(":.*","",chessdata$opening_name))
chessdata<-transform(chessdata, shortname=gsub("\\|.*","",chessdata$shortname))
head(chessdata)
```



# Exploratory charts

Let's take a look at what we have in our new "shortnames" field in terms of number of games for each opening:

```{r listofnames}
length(table(chessdata$shortname))
```

180 is too many openings to neatly chart! Let's filter our table down to the top 15 openings and then take a look at some bar charts.


```{r filter}
top15openings<-head(names(sort(table(chessdata$shortname), decreasing=TRUE)),15)
filtereddata<-filter(chessdata, shortname %in% top15openings)
```

Now let's start with a basic bar chart just to see the relative frequency of these top 15 openings.

```{r basicbarchart}
BarChart(shortname, data=filtereddata, horiz=TRUE, quiet=TRUE, xlab="Opening", sort="-", ylab="Count of Games")
```

The Sicilian Defense (1. e4 c6) is by far the most popular, with an even distribution after that. We're keeping an eye on the Italian Game (1. e4 e5 2. Nf3 Nc6 3. Bc4) since that's my favorite opening, and I'm pleased to see it in 4th place.

As an excuse to switch to histograms, how about the ratings of the players in this data set?

```{r ratinghisto}
Histogram(white_rating, data=filtereddata, fill="lightgray", quiet=TRUE, xlab="White Player's Rating", ylab="Count of Games")
```

This looks like a normal distribution with a slight skew to the left. I'd explain this by saying your default rating is 1500, so there are probably a lot of games by new players at or immediately under 1500. Note the slight uptick at the very top of the range - a buildup of high-rated players.

Are the ratings of the black players any different?
```{r ratinghisto2}
Histogram(black_rating, data=filtereddata, fill="black", quiet=TRUE, xlab="Black Player's Rating", ylab="Count of Games")
```

No, the distribution of ratings for the black players are identical; that's as expected. How long is a normal game?

```{r turnshisto}
Histogram(turns, data=filtereddata, bin.width=15, quiet=TRUE, xlab="Number of Turns", ylab="Count of Games")
```

```{r top4openingratingline}
top4<-head(names(sort(table(chessdata$shortname), decreasing=TRUE)),4)
filtereddata2<-filter(chessdata, shortname %in% top4)
ggplot(filtereddata2, aes(x=white_rating, color=shortname))+geom_line(stat="bin", binwidth=25)+ylim(0,100)
```



```{r winratebarchart}
BarChart(shortname, by=winner, data=filtereddata, horiz=TRUE, ylab="Count of Games", stat.x="proportion", xlab="Opening Name",quiet=TRUE)
```

---most common opening by rating?

--- draw rate by opening and rating

# Best Opening for Me

---highest winrate by rating

```{r winrate by rating}
options(scipen=999)
filtereddata$ratingbin<-cut(filtereddata$white_rating, seq(0,3500,500), dig.lab=10)
##BarChart(ratingbin, by=winner, by1=shortname, data=filtereddata, horiz=TRUE, ylab="Count of Games", stat.x="proportion", xlab="Opening Name",quiet=TRUE)
ggplot(filtereddata, aes(x = ratingbin, fill=winner)) +  
        geom_bar(aes(y = (..count..)/sum(..count..)),position="fill") + 
        scale_y_continuous()+facet_grid(cols=vars(shortname)) ## version 3.0.0
```


---highest winrate by time control

---linechart of performance of top 3 openings by rating

# Shiny Interactive

-- highest winrate by binned rating as a shiny