---
title: "HW1"
author: "Jordan Hilton"
date: "April 2, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lessR)
library(ggplot2)
library(tidyverse)
```


# 1 Bar Chart

Let's load the data:

```{r barchartload}
d <- rd("Employee", format="lessR")
```

## a.

Here's a barplot of the number of employees in each department using the base R plot:

```{r barplot}
barplot(table(d$Dept))
```

## b.

Here's the same data in table form:

```{r table}
table(d$Dept)
```

## c.

Here's the same chart in lessR:

```{r barchart}
BarChart(Dept, quiet=TRUE)
```

and ggplot2:

```{r ggplot2barchart}
ggplot(d, aes(Dept))+geom_bar()
```


## d.

Here's the lessR 1d bubble plot:

```{r bubbleplot}
Plot(Dept, quiet=TRUE)
```

## e.

The bubble chart is more compact and there could be applications where showing relative size as an area as opposed to a length is useful. The bar chart is more readable and more common, so it will make more sense to most readers.

## h. (no f/g?)

Here's the bar chart with proportions instead of counts:

```{r barchart2}
BarChart(Dept, quiet=TRUE, stat.x="proportion")
```

## i.

With horizontal bars:

```{r barchart3}
BarChart(Dept, quiet=TRUE, horiz=TRUE)
```

## j.

Now providing a title and custom axis lables:

```{r barchart4}
BarChart(Dept, quiet=TRUE, xlab="Our Cool Departments", ylab="Awesome Employee Count", main="Employment at our Tubular Company")
```

# 2. R Factors

Let's load the survey data:

```{r surveyload}
surveydata<-rd("460S14.csv", quiet=TRUE)
head(surveydata)
length(surveydata$Learn_1)
```

## a. 

We can see from the length of the first column that there are 31 rows of data, so that's probably the number of student responses we got.

## b.

The four variables we're discussing are Past_1, Past_2, Past_3, and Past_4

## c.

```{r checknas}
sum(is.na(surveydata))
sum(is.na(surveydata$Past_1))
sum(is.na(surveydata$Past_2))
sum(is.na(surveydata$Past_3))
sum(is.na(surveydata$Past_4))
```

We can see that there are 7 total missing values in the dataset, but no missing values for any of these 4 questions. The missing values are in the columns "Past2_1" and "Learn2_1" so I"m guessing that those questions are either optional or only presented under some circumstances.

## d.

Here are the first 3 rows of data for our four variables:

```{r first3}
head(cbind(surveydata$Past_1, surveydata$Past_2, surveydata$Past_3,surveydata$Past_4),3)
```

## e.

It looks like the values are recorded 


